### 2023-12-19 draft_jss/draft_jss.Rmd
update draft

---
documentclass: jss
bibliography: article.bib
author:
  - name: FirstName LastName
    orcid: 0000-0000-0000-0000
    affiliation: University/Company
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{name@company.com}
    url: https://posit.co
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Fourth Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma 
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "Reservoir computing in R : a tutorial using \\pkg{reservoirnet}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Reservoir computing in R : a tutorial using reservoirnet"
  # For running headers, if needed
  short:     "\\pkg{reservoirnet}: reservoir computing in R"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output:
  bookdown::pdf_book:
    base_format: rticles::jss_article
  bookdown::word_document2:
    default
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = here::here("draft_jss"), output_format = "all") })
---

@@

Numerous COVID-19 forecasting algorithms have been proposed using different methods (e.g ensemble, deep learning, compartmental), yet none has proven entirely satisfactory [@rahimi_review_2021, @cramer_evaluation_2022]. In France, short-term forecasts  with different methods have been evaluated with similar results [@pottier_forecast_2021, @paireau_ensemble_2022, @carvalho_analysis_2021, @mohimont_convolutional_2021]. In this context a machine learning algorithm based on linear regression with elastic-net penalization, leveraging both Electronic Health Records (EHRs) and public data, was implemented at Bordeaux University Hospital [@ferte_benefit_2022]. This model, which aimed at forecasting the number of hospitalized patients at 14 days, showed good performance but struggled to accurately anticipate dynamic shifts of the epidemic.

@@

Mean was chosen over median for $MRE$ and $MREB$ because those metrics tend to have extremely high values when the denominator is close to 0 (i.e when the number of hospitalized patients is close to 0 or the number of patients hospitalized at 14 days is close to the current number of hospitalized patients respectively). $MAEB$ and $MREB$ compare model performance to a baseline model which predicts the current number of hospitalized patients at 14 days. Those metrics help to determine the information added by the model and is a good baseline as covid-19 forecast model do not always outperform this basic forecast (@cramer_evaluation_2022).

>>>>

---
documentclass: jss
bibliography: article.bib
author:
  - name: FirstName LastName
    orcid: 0000-0000-0000-0000
    affiliation: University/Company
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{name@company.com}
    url: https://posit.co
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation'
    # To add another line, use \AND at the end of the previous one as above
  - name: Fourth Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma 
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "Reservoir computing in R : a tutorial using \\pkg{reservoirnet}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Reservoir computing in R : a tutorial using reservoirnet"
  # For running headers, if needed
  short:     "\\pkg{reservoirnet}: reservoir computing in R"
abstract: >
  Reservoir Computing (RC) is a machine learning method that uses the dynamics of a high-dimensional "reservoir" to efficiently process information generated by dynamical systems. It has been successful in solving various tasks including time epidemic forecasting, language processing or bird calls. RC is implemented in Python and Julia but not in R. This article introduces **reservoirnet**, a R package providing access to the Python API **ReservoirPy**, allowing R users to harness the power of reservoir computing. This article provides an introduction to the fundamentals of RC and showcases its real-world applicability through three distinct sections. First, we cover the foundational concepts of RC, setting the stage for understanding its capabilities. Next, we delve into the practical usage of **reservoirnet**  through two illustrative examples. These examples demonstrate how it can be applied to real-world problems, specifically, regression of COVID-19 hospitalizations and classification of Japanese vowels. Finally, we present a comprehensive analysis of a real-world application of **reservoirnet**, where it was used to forecast COVID-19 hospitalizations at Bordeaux University Hospital using public data and electronic health records.
keywords:
  # at least one keyword must be supplied
  formatted: ["\\textbf{reservoirnet}", "\\textbf{ReservoirPy}", Reservoir computing, COVID-19, time series]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output:
  bookdown::pdf_book:
    base_format: rticles::jss_article
  bookdown::word_document2:
    default
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = here::here("draft_jss"), output_format = "all") })
---

@@

Numerous COVID-19 forecasting algorithms have been proposed using different methods (e.g ensemble, deep learning, compartmental), yet none has proven entirely satisfactory [@rahimi_review_2021, @cramer_evaluation_2022-1]. In France, short-term forecasts  with different methods have been evaluated with similar results [@pottier_forecast_2021, @paireau_ensemble_2022, @carvalho_analysis_2021, @mohimont_convolutional_2021]. In this context a machine learning algorithm based on linear regression with elastic-net penalization, leveraging both Electronic Health Records (EHRs) and public data, was implemented at Bordeaux University Hospital [@ferte_benefit_2022]. This model, which aimed at forecasting the number of hospitalized patients at 14 days, showed good performance but struggled to accurately anticipate dynamic shifts of the epidemic.

@@

Mean was chosen over median for $MRE$ and $MREB$ because those metrics tend to have extremely high values when the denominator is close to 0 (i.e when the number of hospitalized patients is close to 0 or the number of patients hospitalized at 14 days is close to the current number of hospitalized patients respectively). $MAEB$ and $MREB$ compare model performance to a baseline model which predicts the current number of hospitalized patients at 14 days. Those metrics help to determine the information added by the model and is a good baseline as covid-19 forecast model do not always outperform this basic forecast (@cramer_evaluation_2022-1).
